The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond

### 2019-04-22 paper/ms.tex
Create new section that compiles all Galactocentric stuff

Underlying the quantitative analysis of many of the mechanisms at work that
give rise to these signatures is the assumption of a global, axisymmetric
Galactocentric coordinate system \citep{2008gady.book.....B} in which the
Sun's position (and therefore the origin of the coordinate system) can be
defined and measured both precisely and accurately. In this system, the center
of the Galaxy is usually taken to be the location of the central supermassive
black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
the Galactic center can be measured \citep{2009ApJ...692.1075G,
2018AA...615L..15G}, with a recent measurement using near-infrared
interferometry of $8.178 \pm 0.035$~kpc \citep{2019arXiv190405721A}. The Sun's
vertical distance from the Galactic midplane, the other independent
measurement that sets this coordinate system, can be determined by identifying
where the stellar density and velocities reach a maximum (effectively the
median height of all disk stars), and is usually taken to be $\sim 25\,\pc$
\citep{2001ApJ...553..184C}, with more recent measurements from {\em Gaia} DR2
placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is
to use the cold gas or HII regions in the disk to define the Galactic
midplane, leading to slightly different values ($\sim 5\,\pc$) for the Sun's
relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review
of these measurements is given by \citet{2016ARAA..54..529B}.

@@

\section{Azimuthal midplane variations at the Solar Circle: estimates from
simulation} \label{sec:local_fire}
\subsection{Description of FIRE Simulation} \label{ssec:cosmozoom}
We briefly describe the cosmological-hydrodynamical zoom-in simulations used
in this work. Cosmological zoom-ins allow one to simulate a selected region at
high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called \mi{}, \mf{}, and \mm{}
for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}

@@

Our main point is that the local midplane varies between different points in
the Galaxy. This is a different statement than that our measurements of the
local midplane are erroneous. Indeed, the local midplane has recently been
measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
that determinations of the local midplane are perfect, since midplane
measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
additional complication is the fact that non-zero Solar heights induce a tilt
in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}. We do
not propose here what a modeler should use for the ``true'' global midplane,
since it depends on the particular problem being studied. Even without a
high-precision definition of the global Galactic midplane, current
observations from {\em Gaia} should soon permit a measurement of the real
azimuthal dependence of the midplane location.

>>>>

Underlying the quantitative analysis of many of the mechanisms at work that
give rise to these signatures is the assumption of a global, axisymmetric
Galactocentric coordinate system \citep{2008gady.book.....B} in which the
Sun's position (and therefore the origin of the coordinate system) can be
defined and measured both precisely and accurately.

@@

\section{Coordinate System Offsets I. Measurement Errors} \label{sec:mes_err}

\subsection{Galactic Center}
First, one must define the center of the Galaxy. This is usually taken to be
the location of the central supermassive black hole, Sgr~A\textsuperscript{*}
\citep[e.g.][]{2004ApJ...616..872R}. From stellar motions near
Sgr~A\textsuperscript{*}, the distance from the Sun to the Galactic center can
be measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a recent
measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
\citep{2019arXiv190405721A}.

We take the ``dynamical Galactic center'' to be the point in three-dimensional
space about which the stars in the solar neighborhood are orbiting. An
important point to make about the Galactic center is that while many authors
assume Sgr~A\textsuperscript{*} is colocated with the dynamical Galactic
center, this assumption has not yet been justified. One potential test is to
perform some sort of dynamical modelling where the distance from the Sun to
the Galactic center is a free parameter and can be constrained with real data.
For example, \citet{2015ApJ...803...80K} were able to make this measurement by
modelling the dynamics of the stream Palomar 5. Other dynamical measurements
of the Solar distance have been made, though none with a precision capable of
competing with the distance to Sgr~A\textsuperscript{*}
\citep{1981gask.book.....M, 2011PASJ...63..867S, 2012MNRAS.427..274S,
2013AstL...39...95B, 2013IAUS..289..444Z}.

\subsection{Galactic Orientation}
Second, one must define the orientation of the Galaxy about the center. This
was defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
defining the location of the Galactic center and the north Galactic pole.
However, there is growing evidence that the true midplane is tilted when
considering stars \citep{2014ApJ...797...53G, 2016ARAA..54..529B}, though not
with H~\textsc{ii} \citep{2019ApJ...871..145A}.

\subsection{Solar Height}
Thirdly, one must define The Sun's vertical distance from the Galactic
midplane, which can be determined by identifying where the stellar density
and velocities reach a maximum (effectively the median height of all disk
stars) The Solar height is usually taken to be $\sim 25\,\pc$
\citep{2001ApJ...553..184C}, with a more recent measurement from {\em Gaia} DR2
placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is
to use the cold gas or H~\textsc{ii} regions in the disk to define the Galactic
midplane, leading to slightly different values ($\sim 5\,\pc$) for the Sun's
relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review
of these measurements is given by \citet{2016ARAA..54..529B}.

\subsection{Local Standard of Rest}
Finally, one must define the local standard of rest (LSR). The radial
($U_{\odot}$) and vertical ($W_{\odot}$)components are easily computed by
taking the mean motions of different stellar groups. The azimuthal component
($V_{\odot}$) is more difficult to measure, but can be modelled using the
asymmetric drift relation \citep{2008gady.book.....B}. The values of the
components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
value of the circular velocity is taken to be $\sim 220\kms$
\citep[e.g.][]{2012ApJ...759..131B}.

In order to convert a star's position and velocity from heliocentric to
Galactocentric coordinates, a number of parameters must be measured. Given the
great interest in each parameter, considerable effort has been placed on these
measurements. However, uncertainties remain, and detailed dynamical modelling
across the disk --- particularly for dynamically cold stars --- may have to
take them into account. For instance, the discrepancy of $\sim 5\,\pc$ between
the midplane defined by stars and by gas will induce a $\sim10\%$ error in
$J_z$ for an orbit with $z_{\text{max}}\sim100\,\pc$ (see
Section~\ref{sec:ref_frame}).

% \section{Azimuthal midplane variations at the Solar Circle: estimates from
% simulation} 
\section{Coordinate System Offsets II. Azimuthal Midplane Variations in
Simulations}
\label{sec:local_fire}
\subsection{Description of FIRE Simulation} \label{ssec:cosmozoom}
We briefly describe the cosmological-hydrodynamical zoom-in simulations used
in this work. Cosmological zoom-ins allow one to simulate a selected region at
high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called \mi{}, \mf{}, and \mm{}
for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}

@@

Our main point is that the local midplane varies between different points in
the Galaxy. We do
not propose here what a modeler should use for the ``true'' global midplane,
since it depends on the particular problem being studied. Even without a
high-precision definition of the global Galactic midplane, current
observations from {\em Gaia} should soon permit a measurement of the real
azimuthal dependence of the midplane location.
